---
title: "Analysis Part 1: SENSITIVITY"
subtitle: "Descriptive Results of Similarity Ratings"
date: "`r Sys.Date()`"
author: "David Izydorczyk"
format: 
  html:
    code-fold: true
    code-copy: true
    code-tools: true
    df-print: kable
    toc: true
    toc-location: right
    fig-height: 4
    fig-width:  9
    fig-cap-location: top
    tab-cap-location: top
    embed-resources: true
    grid:
      margin-width: 350px
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| code-summary: "Load packages & ggplot theme"
#| cache: false

# Packages
library(tidyverse)
library(psych)
library(purrr)
library(patchwork)
library(knitr)
library(kableExtra)
library(stringr)
library(tm)
library(ggwordcloud)
library(wordcloud)
library(viridis)
library(correlation)



source("../../Scripts/plot_settings.R")
source("../../Scripts/helper_functions.R")
```

```{r data}
#| code-summary: "Load data"

# Load data sets

df_food    <- read_csv2("../../Data/Sensitivity Checks/data_analysis_similarity_food_SENSITIVITY.csv") |>
                  rename(img_stim_1 = stim_1, img_stim_2 = stim_2,stim_1=food_stim_1,stim_2=food_stim_2)

df_country <- read_csv2("../../Data/Sensitivity Checks/data_analysis_similarity_country_SENSITIVITY") |>
                  rename(img_stim_1 = stim_1, img_stim_2 = stim_2,stim_1=country_stim_1,stim_2=country_stim_2) 

df_mammal  <- read_csv2("../../Data/Sensitivity Checks/data_analysis_similarity_mammals_SENSITIVITY.csv") |> 
                  rename(img_stim_1 = stim_1, img_stim_2 = stim_2,stim_1=mammal_stim_1,stim_2=mammal_stim_2) 


# Load stimulus List

stims_food    <- read_csv2("../../Materials/stimlist_foods.csv")
stims_country <- read_csv2("../../Materials/stimlist_countries.csv") |> rename(category = continent)
stims_mammal  <- read_csv2("../../Materials/stimlist_mammals.csv") 


# Add category/continent/family infos to dfs

df_food    <- left_join(df_food,stims_food[,c("ID","category")], by =  join_by(ID_stim_1  == ID)) |> 
                left_join(stims_food[,c("ID","category")], by =  join_by(ID_stim_2  == ID),suffix = c("_1", "_2"))

df_country <- left_join(df_country, stims_country[,c("ID","category")], by =  join_by(ID_stim_1  == ID)) |> 
                left_join(stims_country[,c("ID","category")], by =  join_by(ID_stim_2  == ID),suffix = c("_1", "_2"))

df_mammal <- left_join(df_mammal, stims_mammal[,c("ID","category")], by =  join_by(ID_stim_1  == ID)) |> 
                left_join(stims_mammal[,c("ID","category")], by =  join_by(ID_stim_2  == ID),suffix = c("_1", "_2"))

# Combine into tibble
dfs <- tibble(names   = c("Food", "Countries", "Mammals"),
              dataset = list(df_food,df_country,df_mammal))

```


# Descriptive Statistics                                 

```{r}
#| code-summary: "Compute descriptive statistics of similarity ratings"


# Compute average similarities for all pairs
dfs <- dfs |> 
          mutate(sim_mat = purrr::map(dataset, ~ .x |> 
                            group_by(ID_stim_1, ID_stim_2, stim_1, stim_2, category_1, category_2) |> 
                            summarise(norm_sim = mean(norm_sim), .groups = "drop") |> 
                            select(ID_stim_1, ID_stim_2, norm_sim, stim_1, stim_2, category_1, category_2)))



# Compute descriptives per domain
sim_summary <- dfs  |> 
                transmute(names, sim_stats = purrr::map(sim_mat, ~ .x |> pull(norm_sim) |> 
                                                           describe() |> as_tibble()))  |> 
                unnest(sim_stats)


sim_summary |>
  select(names, mean, sd, min, max) |>
  kable(digits = 2)
```

# Pairwise Similarity Matrices                           

```{r}
#| code-summary: "Make Figure 3 (Pairwise Similarities)"
#| fig-height: 8
#| fig-width:  20
#| fig-align: "center"
#| label: fig-sim
#| fig-cap: "Average Pairwise similarites"
#| column: page

# Reorder stimuli based on categories
order_food <- dfs$sim_mat[[1]] |> 
                bind_rows(data.frame(ID_stim_1 = 1, 
                                     ID_stim_2 = 80,
                                     norm_sim  = 0.09444444,
                                     stim_1    = "Traubenzucker",
                                     stim_2    = "Traubenzucker",
                                     category_1 = "Sweets and snacks", 
                                     category_2 = "Sweets and snacks")) |> # To make Traubenzucker appear in the data
                arrange(category_1,category_2,stim_1,stim_2) |> 
                transmute(plot_ID   = as.integer(factor(stim_1, levels = unique(stim_1))),
                          plot_stim = stim_1) |> 
                distinct()

# Category order:

# [1] "Cereal products and potatoes" "Dairy products and eggs"     
# [3] "Fruits"                       "Legumes"                     
# [5] "Meat (substitutes) and fish"  "Nuts"                        
# [7] "Spreads"                      "Sweets and snacks"           
# [9] "Vegetables"

temp_food <- dfs$sim_mat[[1]] |> 
                left_join(order_food, by =  join_by(stim_1  == plot_stim), relationship = "many-to-many") |> 
                left_join(order_food, by =  join_by(stim_2  == plot_stim), relationship = "many-to-many",suffix = c("_1", "_2"))
   

temp_food |> 
  mutate(plot_ID_min = pmin(plot_ID_1, plot_ID_2),
         plot_ID_max = pmax(plot_ID_1, plot_ID_2)) |> 
  arrange(plot_ID_min, plot_ID_max) |> 
  ggplot(aes(x = plot_ID_max, y = plot_ID_min, fill = norm_sim)) +
      geom_tile() +
      labs(x = expression("Stim " * italic(i)),
           y = expression("Stim " * italic(j)),
           fill = "Normalized Mean Similarities", title = "Foods") +
      scale_x_continuous(expand = c(0.01, 0),breaks = c(1,seq(from = 0, to = 80, by = 5))) + 
      scale_y_continuous(expand = c(0.01, 0),breaks = c(1,seq(from = 0, to = 80, by = 5)),limits = c(0,100)) + 
      scale_fill_viridis_b() +
      theme_nice() +
      theme(legend.position = "right") + 
      ggplot2::annotate("text", x = 5, y = 20, label = "Cereal products \n & potatoes", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 17, y = 30, label = "Dairy products \n & eggs", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 29, y = 35, label = "Fruits", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 37, y = 47, label = "Legumes", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 45, y = 59, label = "Meat (substitutes) \n & fish", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 51, y = 59, label = "Nuts", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 55, y = 64, label = "Spreads", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 61, y = 70, label = "Sweets & \n snacks", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 75, y = 85, label = "Vegetables", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      theme(axis.ticks.x = element_blank(),
            axis.ticks.y = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            panel.grid.major = element_blank()) -> sim_mat_plot_food


# Reorder stimuli based on categories
order_countries <- dfs$sim_mat[[2]] |> 
                bind_rows(data.frame(ID_stim_1 = 1, 
                                     ID_stim_2 = 80,
                                     norm_sim  = 99999,
                                     stim_1    = "Vatican City",
                                     stim_2    = "Vatican City",
                                     category_1 = "europe", 
                                     category_2 = "europe")) |> # To make Traubenzucker appear in the data
                arrange(category_1,category_2,stim_1,stim_2) |> 
                transmute(plot_ID   = as.integer(factor(stim_1, levels = unique(stim_1))),
                          plot_stim = stim_1) |> 
                distinct()

# Category order: 
# order_countries$category_1 |> unique()
# [1] "africa"        "asia"          "europe"        "north_america"
# [5] "oceania"       "south_america"

temp_countries <- dfs$sim_mat[[2]] |> 
                     left_join(order_countries, by =  join_by(stim_1  == plot_stim), relationship = "many-to-many") |> 
                     left_join(order_countries, by =  join_by(stim_2  == plot_stim), relationship = "many-to-many",
                               suffix = c("_1", "_2"))
   

temp_countries |> 
  mutate(plot_ID_min = pmin(plot_ID_1, plot_ID_2),
         plot_ID_max = pmax(plot_ID_1, plot_ID_2)) |> 
  arrange(plot_ID_min, plot_ID_max) |> 
  ggplot(aes(x = plot_ID_max, y = plot_ID_min, fill = norm_sim)) +
      geom_tile() +
      labs(x = expression("Stim " * italic(i)),
           y = "",
           fill = "Normalized Mean Similarities", title = "Countries") +
      scale_x_continuous(expand = c(0.01, 0),breaks = c(1,seq(from = 0, to = 80, by = 5))) + 
      scale_y_continuous(expand = c(0.01, 0),breaks = c(1,seq(from = 0, to = 80, by = 5)),limits = c(0,100)) + 
      scale_fill_viridis_b() +
      theme_nice() +
      theme(legend.position = "right") + 
      ggplot2::annotate("text", x = 7, y = 15, label = "Africa", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 22, y = 30, label = "Asia", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 36, y = 45, label = "Europe", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) +
      ggplot2::annotate("text", x = 48, y = 60, label = "North & Central \n America", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) + 
      ggplot2::annotate("text", x = 60, y = 70, label = "Oceania", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) + 
      ggplot2::annotate("text", x = 74, y = 87, label = "South America", 
                        size = 5, vjust = 0.5 , angle = 90, lineheight = 0.8) + 
      theme(axis.ticks.x = element_blank(),
            axis.ticks.y = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            panel.grid.major = element_blank())   -> sim_mat_plot_countries



# Reorder stimuli based on categories
order_mammals <- dfs$sim_mat[[3]] |> 
                bind_rows(data.frame(ID_stim_1 = 1, 
                                     ID_stim_2 = 80,
                                     norm_sim  = 99999,
                                     stim_1    = "Asiatischer Elefant",
                                     stim_2    = "Asiatischer Elefant",
                                     category_1 = "Proboscidea", 
                                     category_2 = "Proboscidea")) |>
                arrange(category_1,category_2,stim_1,stim_2) |> 
                transmute(plot_ID   = as.integer(factor(stim_1, levels = unique(stim_1))),
                          plot_stim = stim_1, category_1) |> 
                distinct()

# Category order: 
# order_mammals$category_1 |> unique()
#  [1] "Artiodactyla"   "Carnivora"      "Cetacea"        "Dasyuromorphia" "Diprotodontia" 
#  [6] "Lagomorpha"     "Mammalia"       "Monotremata"    "Perissodactyla" "Pilosa"        
# [11] "Primates"       "Proboscidea"    "Rodentia"       "Sirenia"     

temp_mammals <- dfs$sim_mat[[3]] |> 
                     left_join(order_mammals, by =  join_by(stim_1  == plot_stim), relationship = "many-to-many") |> 
                     left_join(order_mammals, by =  join_by(stim_2  == plot_stim), relationship = "many-to-many",
                               suffix = c("_1", "_2"))
   

temp_mammals |> 
  mutate(plot_ID_min = pmin(plot_ID_1, plot_ID_2),
         plot_ID_max = pmax(plot_ID_1, plot_ID_2)) |> 
  arrange(plot_ID_min, plot_ID_max) |> 
  ggplot(aes(x = plot_ID_max, y = plot_ID_min, fill = norm_sim)) +
      geom_tile() +
      labs(x = expression("Stim " * italic(i)),
           y = "",
           fill = "Normalized Mean Similarities", title = "Mammals") +
      scale_x_continuous(expand = c(0.01, 0),breaks = c(1,seq(from = 0, to = 80, by = 5))) + 
      scale_y_continuous(expand = c(0.01, 0),breaks = c(1,seq(from = 0, to = 80, by = 5)),limits = c(0,100)) + 
      scale_fill_viridis_b() +
      theme_nice() +
      theme(legend.position = "right") + 
      ggplot2::annotate("text", x = 7, y = 19, label = "Artiodactyla", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) +
      ggplot2::annotate("text", x = 24, y = 35, label = "Carnivora", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) +
      ggplot2::annotate("text", x = 34.5, y = 43, label = "Cetacea", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) +
      ggplot2::annotate("text", x = 37, y = 52, label = "Dasyuromorphia", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 40.5, y = 57, label = "Diprotodontia", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 43, y = 60, label = "Lagomorpha", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 45, y = 62, label = "Mammalia", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 47, y = 64, label = "Monotremata", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 49.5, y = 68, label = "Perissodactyla", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 52, y = 69, label = "Pilosa", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 61, y = 75, label = "Primates", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 70.5, y = 82, label = "Proboscidea", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 76, y = 85, label = "Rodentia", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      ggplot2::annotate("text", x = 79, y = 87, label = "Sirenia", 
                        size = 5, vjust = 0.5, angle = 90, lineheight = 0.9) + 
      theme(axis.ticks.x = element_blank(),
            axis.ticks.y = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            panel.grid.major = element_blank())  -> sim_mat_plot_mammals



sim_mat_plot_food + sim_mat_plot_countries + sim_mat_plot_mammals &
  plot_layout(guides = "collect") +
  theme(legend.position = 'bottom',
        legend.key.width=unit(1,"cm")) + 
  scale_fill_continuous(limits = c(0,1),type="viridis") +
  guides(fill = guide_colorbar(title = "Mean Normalized Similarity    ",
                               label.position = "bottom",
                               title.position = "left", title.vjust = 1,
                               # draw border around the legend
                               frame.colour = "black",
                               barwidth = 15,
                               barheight = 1.5)) 

```

# Table with 5 most similar and dissimilar countries     

@tbl-sim shows the on average 10 most similar and 10 most dissimilar countries.
 
```{r}
#| code-summary: "Make tables"
#| label: tbl-sim
#| tbl-cap: "5 most similar and dissimilar countries (on average)"
#| tbl-subcap: 
#| - "Food"
#| - "Countries"
#| - "Mammals"
#| layout-ncol: 3
#| column: page


create_latex_table <- function(df, dataset_name) {
  
  df |> 
    select(stim_1,stim_2,norm_sim) |> 
    arrange(norm_sim) |> 
    mutate(norm_sim = round(norm_sim, 2)) |> 
    slice(c(1:5, (n() - 4):n())) |> 
    mutate(across(everything(), as.character)) %>%  # ensure all are character
    kbl(col.names = c("Stimulus i", "Stimulus j", "M"),align = "c") |> 
    kable_paper(c("hover")) 
  
}



# Apply to all datasets in dfs
dfs <- dfs |> mutate(tbl = map2(sim_mat, names, create_latex_table))

dfs$tbl[[1]] 
dfs$tbl[[2]]
dfs$tbl[[3]]
```


# Descriptive: Features                                  

@fig-wc gives an overview of the features people used to make their similarity judgments. Raw responses where preprocessed (lower case, remove punctuation and white space), lemmatized (i.e., "culture"/"cultural" -> "culture"). Only feature which were reported at least 3 times are plotted.


```{r}
#| code-summary: "Plot Wordcloud"
#| fig-height: 6
#| fig-width:  14
#| fig-align: "center"
#| label: fig-wc
#| fig-cap: "Wordcloud of reported features"
#| column: page


preprocessing <- function(df){
  
  features <- df %>% select(ID, features) %>% distinct()

  # Clean and preprocess the text
  clean_text <- features$features
  
  # Create a text corpus 
  corpus <- Corpus(VectorSource(clean_text))
  
  # Convert to lower case
  corpus <- tm_map(corpus, content_transformer(tolower))
  
    # Remove extra whitespace
  corpus <- tm_map(corpus, stripWhitespace)
  
  # Remove punctuation
  corpus <- tm_map(corpus, removePunctuation)
  
  # Remove stopwords (common words that don’t add much meaning)
  corpus <- tm_map(corpus, removeWords, stopwords("en"))
  
  # Remove other filler words
  words2remove <- c("answer","etc","say","try","two","come","give","similar",
                    "similarity","food","country","animal", "sometimes", "whether",
                    "can","mostly","first","slightly","alike","give","even","start",
                    "relate","similarity","seem","just","another","somewhat","theyre",
                    "may","mostly","one","just","feature","didnt","way","guess")
  
  corpus  <- tm_map(corpus, removeWords, words2remove)
  

  # Optionally, remove numbers
  corpus <- tm_map(corpus, removeNumbers)
  
  corpus <- tm_map(corpus, content_transformer(function(x) textstem::lemmatize_strings(x)))
  
  # Create a term-document matrix
  tdm <- TermDocumentMatrix(corpus)
  
  return(tdm)
  
}

plot_wc <- function(tdm, name){
  
  # Get word frequencies
  word_freqs      <- sort(rowSums(as.matrix(tdm)), decreasing = TRUE)
  word_freq_table <- data.frame(word = names(word_freqs), freq = word_freqs)
  
  # Generate the word cloud again
  ggwordcloud(words    = word_freq_table$word,
            freq     = word_freq_table$freq,
            min.freq = 5, 
            scale    = c(4,0.5), 
            colors   = viridis(8,option = "D"),
            family = "Jost") + 
    labs(title = name) + 
    theme(plot.title   = element_text(hjust = 0.5,size = 20, face = "bold"),
          panel.border = element_rect(colour = "black", linewidth = 0.5, fill = NA))
  
}


dfs %>%
  mutate(tdm     = purrr::map(dataset, preprocessing),
         wc_plot = map2(tdm,names,plot_wc)) |> 
  pull(wc_plot) |> 
  wrap_plots()


```


# Compare Similarity vs Similarity Sensitivity           

Results below shows the correspondence between the average similarity ratings of each pair based on the inclusion criteria used in the main manuscript and the sensitivity analysis without exclusion based on RTs and *r*. 


```{r}
#| code-summary: "Correspondence (Food)"

d_sensitiviy <- read_csv2("../../Results/Sensitivity Checks/dist_mat_food.csv") |> unlist() |> na.omit()
d_normal     <- read_csv2("../../Results/dist_mat_food.csv") |> unlist() |>  na.omit()


d_food <- data.frame(d_normal,d_sensitiviy)
correlation(d_food)

```

```{r}
#| code-summary: "Correspondence (Countries)"

d_sensitiviy <- read_csv2("../../Results/Sensitivity Checks/dist_mat_countries.csv") |> unlist() |> na.omit()
d_normal     <- read_csv2("../../Results/dist_mat_countries.csv") |> unlist() |>  na.omit()


d_countries <- data.frame(d_normal,d_sensitiviy)
correlation(d_countries)

```

```{r}
#| code-summary: "Correspondence (Mammals)"

d_sensitiviy <- read_csv2("../../Results/Sensitivity Checks/dist_mat_mammals.csv") |> unlist() |> na.omit()
d_normal     <- read_csv2("../../Results/dist_mat_mammals.csv") |> unlist() |>  na.omit()


d_countries <- data.frame(d_normal,d_sensitiviy)
correlation(d_countries)

```
