---
title: "Analysis Part 2:"
subtitle: "Multidimensional Scaling"
date: "`r Sys.Date()`"
author: "David Izydorczyk"
format: 
  html:
    code-fold: true
    code-copy: true
    code-tools: true
    df-print: kable
    toc: true
    toc-location: right
    fig-height: 4
    fig-width:  9
    fig-cap-location: top
    tab-cap-location: top
    embed-resources: true
    grid:
      margin-width: 350px
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| code-summary: "Load packages & ggplot theme"
#| cache: false

# Packages
library(tidyverse)
library(MDShelper)
library(smacof)
library(purrr)
library(patchwork)
library(knitr)
library(kableExtra)
library(correlation)
library(papaja)
library(ggrepel)
library(ggimage)

source("../Scripts/plot_settings.R")
```

```{r data}
#| code-summary: "Load data"

# Load data sets

df_food    <- read_csv2("../Data/3 Food/data_analysis_similarity_food.csv") |>
                  rename(img_stim_1 = stim_1, img_stim_2 = stim_2,stim_1=food_stim_1,stim_2=food_stim_2)

df_country <- read_csv2("../Data/1 Countries/data_analysis_similarity_country.csv") |>
                  rename(img_stim_1 = stim_1, img_stim_2 = stim_2,stim_1=country_stim_1,stim_2=country_stim_2) 

df_mammal  <- read_csv2("../Data/2 Mammals/data_analysis_similarity_mammals.csv") |> 
                  rename(img_stim_1 = stim_1, img_stim_2 = stim_2,stim_1=mammal_stim_1,stim_2=mammal_stim_2) 


# Load item information (with image paths)
stims_food    <- read_csv2("../Materials/stimlist_foods.csv")
stims_country <- read_csv2("../Materials/stimlist_countries.csv") |> rename(category = continent)
stims_mammal  <- read_csv2("../Materials/stimlist_mammals.csv") 

# Combine into tibble
dfs <- tibble(names   = c("Food", "Countries", "Mammals"),
              dataset = list(df_food,df_country,df_mammal),
              stims   = list(stims_food,stims_country,stims_mammal))

```

```{r}
#| code-summary: "Create pairwise distance matrices"

prepare_data <- function(df){
  
  df |> 
    select(ID,ID_stim_1,ID_stim_2,norm_dist) %>%   
    group_by(ID_stim_1,ID_stim_2) %>% 
    summarize(avg_dist = mean(norm_dist,na.rm=T),.groups="drop") %>%  
    arrange(ID_stim_1,ID_stim_2) %>%  
    pivot_wider(names_from = ID_stim_1,values_from = avg_dist) %>% 
    .[,-1] %>%  
    as.matrix() %>%  
    rbind(NA,.) %>%  
    cbind(.,NA)

}

dfs <- dfs |>  mutate(dist_mat = purrr::map(dataset,prepare_data))

# Save distance matrices
write_csv2(dfs$dist_mat[[1]] |> as.data.frame(),file="../Results/dist_mat_food.csv")
write_csv2(dfs$dist_mat[[2]] |> as.data.frame(),file="../Results/dist_mat_countries.csv")
write_csv2(dfs$dist_mat[[3]] |> as.data.frame(),file="../Results/dist_mat_mammals.csv")

```


# Multidimensional Scaling                                

```{r}
#| code-summary: "Run loo-CV"

# UNCOMMENT TO RUN LOO (~ 1h per data set)
# loo_CV_MDS(dfs$dist_mat[[1]],
#            max_dim  = 20,
#            parallel = TRUE,
#            n_cores  = 8,
#            verbose  = TRUE) -> mds_loo_cv_FOOD

# save(mds_loo_cv_FOOD,file="../Results/mds_loo_cv_food.Rdata")


# loo_CV_MDS(dfs$dist_mat[[2]],
#            max_dim  = 20,
#            parallel = TRUE,
#            n_cores  = 8,
#            verbose  = TRUE) -> mds_loo_cv_COUNTRIES


# save(mds_loo_cv_FOOD,file="../Results/mds_loo_cv_countries.Rdata")


# loo_CV_MDS(dfs$dist_mat[[3]],
#            max_dim  = 20,
#            parallel = TRUE,
#            n_cores  = 8,
#            verbose  = TRUE) -> mds_loo_cv_MAMMALS


# save(mds_loo_cv_FOOD,file="../Results/mds_loo_cv_mammals.Rdata")


mds_loo_cv_FOOD      <- get(load("../Results/mds_loo_cv_food.Rdata"))
mds_loo_cv_COUNTRIES <- get(load("../Results/mds_loo_cv_countries.Rdata"))
mds_loo_cv_MAMMALS   <- get(load("../Results/mds_loo_cv_mammals.Rdata"))


# Put in dfs tibble
dfs$mds_loo_cv <- list(mds_loo_cv_FOOD,mds_loo_cv_COUNTRIES,mds_loo_cv_MAMMALS)

```

```{r}
#| code-summary: "Results"
#| label: tbl-mdscv
#| tbl-cap: "Results of MDS Cross-Validation"
#| tbl-subcap: 
#| - "Food"
#| - "Countries"
#| - "Mammals"
#| layout-ncol: 3

mds_res_tbl <- function(results) {
  
  results |>  
    round(3) |>  
    kbl(col.names = c("# Dim", "R²", "RMSE", "MSE"),align = "c") |> 
    kable_paper(c("hover")) |>  
    row_spec(which.min(results[,3]), bold = T, color = "white", background = clrs[2]) |>  
    column_spec(1, width="6em")
}



  

# Apply to all datasets in dfs
dfs <- dfs |> mutate(mds_res = purrr::map(mds_loo_cv, mds_res_tbl))

dfs$mds_res[[1]] 
dfs$mds_res[[2]]
dfs$mds_res[[3]]
```

# Determine MDS Representations (based on loo-cv)         


```{r}
#| code-summary: "Run MDS"
# Run MDS

# MDS_FOOD      <- dfs$dist_mat[[1]] %>%
#                     MDShelper::fill_mat(diag_val = 0) %>%
#                     smacof::mds(.,ndim=14, "ordinal",  principal = TRUE)

# write_csv2(MDS_FOOD$conf %>% as.data.frame(), file="../Results/MDS_config_food.csv")

# MDS_COUNTRIES <- dfs$dist_mat[[2]] %>%
#                     MDShelper::fill_mat(diag_val = 0) %>%
#                     smacof::mds(.,ndim=10, "ordinal",  principal = TRUE)

# write_csv2(MDS_COUNTRIES$conf %>% as.data.frame(), file="../Results/MDS_config_countries.csv")

# MDS_MAMMALS   <- dfs$dist_mat[[3]] %>%
#                     MDShelper::fill_mat(diag_val = 0) %>%
#                     smacof::mds(.,ndim=10, "ordinal",  principal = TRUE)

# write_csv2(MDS_MAMMALS$conf %>% as.data.frame(), file="../Results/MDS_config_mammals.csv")


# Load

MDS_FOOD      <- read_csv2("../Results/MDS_config_food.csv")
MDS_COUNTRIES <- read_csv2("../Results/MDS_config_countries.csv")
MDS_MAMMALS   <- read_csv2("../Results/MDS_config_mammals.csv")


# Put in dfs tibble
dfs$mds_repr <- list(MDS_FOOD,MDS_COUNTRIES,MDS_MAMMALS)


```


```{r}
#| code-summary: "Run MDS with D=10"

# n_dim <- 20
# 
# MDS_FOOD      <- dfs$dist_mat[[1]] %>%
#                     MDShelper::fill_mat(diag_val = 0) %>%
#                     smacof::mds(.,ndim=n_dim, "ordinal",  principal = TRUE)
# 
# write_csv2(MDS_FOOD$conf %>% as.data.frame(col.names=paste0("dim_",1:20)),
#            file="../Results/MDS_config_food_D20.csv")
# 
# MDS_COUNTRIES <- dfs$dist_mat[[2]] %>%
#                     MDShelper::fill_mat(diag_val = 0) %>%
#                     smacof::mds(.,ndim=n_dim, "ordinal",  principal = TRUE)
# 
# write_csv2(MDS_COUNTRIES$conf %>% as.data.frame(col.names=paste0("dim_",1:20)),
#            file="../Results/MDS_config_countries_D20.csv")
# 
# MDS_MAMMALS   <- dfs$dist_mat[[3]] %>%
#                     MDShelper::fill_mat(diag_val = 0) %>%
#                     smacof::mds(.,ndim=n_dim, "ordinal",  principal = TRUE)
# 
# write_csv2(MDS_MAMMALS$conf %>% as.data.frame(col.names=paste0("dim_",1:20)),
#            file="../Results/MDS_config_mammals_D20.csv")

```


# Figure: Scatterplot of observed vs. predicted distances 

@fig-obsemp shows both the predicted distances of this final MDS-solution and the observed distances, as well as their correlation.

```{r}
#| code-summary: "Run MDS + Correlation + Plot"
#| label: fig-obsemp
#| fig-cap: "Scatterplot of observed and predicted distances"
#| fig-height: 6
#| fig-width:  15
#| column: page


obs_pred_plot <- function(mds_repr,dist_mat,names){
  
  # Compute Correlation
  predicted <- mds_repr %>% as.matrix() %>% MDShelper::compute_dists(.,r=2) %>% as.vector() %>% na.omit()
  observed  <- dist_mat %>% as.vector() %>%  na.omit()
  cor       <- correlation(data.frame(predicted,observed))
  
  # Plot Observed ~ Predicted
  p <- ggplot(data.frame(predicted,observed), aes(observed,predicted)) +
    geom_point(shape=1, color = clrs[2],  size = 1.5, alpha = 0.35, stroke = 2) +
    stat_smooth(method="lm",color="black",lwd=1.5) + 
    ggplot2::annotate("text", x = 0.25, y = .75, parse=TRUE,
                      label = paste0("italic(r)==",printnum(cor$r,gt1=F)), size = 7) +
    theme_nice() +
    labs(title = names,
         y = ifelse(names == "Food","Predicted Distances",""),
         x = ifelse(names == "Countries","Observed Distances",""))
  
  return(p)

}


dfs|> 
  mutate(op_plot = pmap(list(mds_repr,dist_mat,names),obs_pred_plot)) |> 
  pull(op_plot) |> 
  wrap_plots()


ggsave("../Figures/oberserved_predicted_dists.pdf", width = 40 , height = 15 , unit = "cm",
       bg="white",dpi=300,device = cairo_pdf)
```

# Figure: Plot of first four dimensions                  

@fig-mdsdim shows the combinations of the four MDS dimensions (**plotted with `check_overlap = TRUE`**).

```{r}
#| code-summary: "Plot MDS Dimensions"
#| label: fig-mdsdim
#| fig-cap: "Dimensions of the MDS Solution"
#| fig-height: 7
#| fig-width: 12
#| column: page

dfs$img_path <- c("../Materials/images/Reduce Quality/Food/",
                  "../Materials/images/Reduce Quality/Countries/",
                  "../Materials/images/Reduce Quality/Mammals/")


plot_df <- function(mds_repr,stims,img_path){
  
  # Make tidy df for plotting
  temp <- stims |>
            select(ID,item,img) |> 
            bind_cols(mds_repr) |> 
            mutate(img_w_path = paste0(img_path,img))
  
  return(temp)
  
}
  

dfs <- dfs |> mutate(plot_df = pmap(list(mds_repr,stims,img_path),plot_df))




# Plot Dimensions 1-2, 3-4 for each domain

# Food
p12_f <- ggplot(dfs$plot_df[[1]], aes(x = V1, y = V2, image=img_w_path)) +
            geom_image(size=0.15) +
            labs(x = "", y = "2. Dimension", title = dfs$names[[1]]) +
            theme_nice() +
            theme(text = element_text(size  = 16)) +
            xlim(-0.5,0.6) + ylim(-0.6,0.45)
 
p34_f <- ggplot(dfs$plot_df[[1]], aes(x = V3, y = V4, image=img_w_path)) +
            geom_image(size=0.15) +
            labs(x = "", y = "4. Dimension") +
            theme_nice() +
            theme(text = element_text(size  = 16)) +
            xlim(-0.5,0.65) + ylim(-0.45,0.6)


# Countries
p12_c <- ggplot(dfs$plot_df[[2]], aes(x = V1, y = V2, image=img_w_path)) +
            geom_image(size=0.23) +
            labs(x = "1. Dimension", y = "", title = dfs$names[[2]]) +
            theme_nice() +
            theme(text = element_text(size  = 16)) +
            xlim(-0.8,0.6) + ylim(-0.5,0.65)
 
p34_c <- ggplot(dfs$plot_df[[2]], aes(x = V3, y = V4, image=img_w_path)) +
            geom_image(size=0.23) +
            labs(x = "3. Dimension", y = "") +
            theme_nice() +
            theme(text = element_text(size  = 16)) +
            xlim(-0.63,0.63) + ylim(-0.5,0.55)
 
 

# Mammals
p12_m <- ggplot(dfs$plot_df[[3]], aes(x = V1, y = V2, image=img_w_path)) +
            geom_image(size=0.15) +
            labs(x = "", y = "", title = dfs$names[[3]]) +
            theme_nice() +
            theme(text = element_text(size  = 16)) +
            xlim(-0.65,0.62) + ylim(-0.75,0.55)
 
p34_m <- ggplot(dfs$plot_df[[3]], aes(x = V3, y = V4, image=img_w_path)) +
            geom_image(size=0.15) +
            labs(x = "", y = "") +
            theme_nice() +
            theme(text = element_text(size  = 16)) +
            xlim(-0.91,0.4) + ylim(-0.35,0.65)




((p12_f + p12_c + p12_m) + theme(plot.margin = unit(c(0,0,30,0), "pt"))) / (p34_f + p34_c + p34_m)


ggsave("../Figures/example_dimensions.pdf", width = 32 , height = 20 , unit = "cm",
       bg="white",dpi=300,device = cairo_pdf)

ggsave("../Figures/example_dimensions.png", width = 32 , height = 20 , unit = "cm",
       bg="white",dpi=300)
```

